Nature Index product

### 2023-09-15 medianSummerTemperature.Rmd
conditional path

# Median Summer Temperature {#median-summer-temp}

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
library(tictoc)
library(ncmeta)
library(parallel)
library(exactextractr)
library(data.table)
library(terra)
dir <- substr(getwd(), 1,2)
pData <- ifelse(dir == "C:", 
  "P:/41201785_okologisk_tilstand_2022_2023/data/",
  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/")
```

```{r, echo=F}
Ecosystem <- "All" 
Egenskap  <- "Abiotiske egenskaper"
ECT       <- "Physical state characteristics" 
Contact   <- "Anders L. Kolstad" 
```

@@

```{r, warning=F}
reg <- st_intersection(reg, nor)
```

```{r, fig.cap="Five accounting areas (regions) in Norway."}
tm_shape(reg) +
  tm_polygons(col="region")

@@

This takes about 20 sec per file/year, or 22 min on total. That is not too bad. About 6000 rasters are read into memory. Here's a test for the effect of splitting over more cores.

```{r, eval=F}
write_stars(real_temp_summer, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/real_temp_summer.tiff")
```

```{r, include=F}
real_temp_summer <- read_stars(paste0(pData, "climate_indicators/aggregated_climate_time_series/real_temp_summer.tiff"),
                                 proxy=F)
```

@@

```{r}
summer_median_temp <- real_temp_summer %>% 
  st_set_dimensions(names = c("x", "y", "v_YEAR")) %>%
  setNames("temperature")
```

```{r, fig.cap="Showing three random slices of the year dimension."}
ggplot()+
  geom_stars(data = summer_median_temp[,,,c(1,11,66)], downsample = c(10, 10, 0)) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))+
  facet_wrap(~v_YEAR)
```

@@

```{r}
median_sd <- function(x) { c(median = median(x), sd = sd(x))}
```

@@

```{r, eval=F}
write_stars(summer_median_temp_ref, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp_ref.tiff")
```

```{r, eval=T, include=F}
summer_median_temp_ref <- read_stars(paste0(pData, "climate_indicators/aggregated_climate_time_series/summer_median_temp_ref.tiff"),
                                     proxy=F)
```

Pivot and turn dimension into attributes, and rename attributes:

```{r, eval=T}
summer_median_temp_ref_long <- summer_median_temp_ref %>% 
  split("band") %>%
  setNames(c("reference_upper", "sd"))
```

@@

```{r}
y_var <- summer_median_temp %>%
  split("v_YEAR") %>%
  c(summer_median_temp_ref_long)
```

@@

```{r, eval=F}
gc()
saveRDS(y_var_norm, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp_normalised.RData")

# Tiff dont allow for multiple attributes:
#write_stars(y_var_norm, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp_normalised.tiff")
```

```{r, include=FALSE}
y_var_norm <- readRDS(paste0(pData, "climate_indicators/aggregated_climate_time_series/temp_median_summer_normalised.RData"))
```

>>>>

# Median Summer Temperature

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse) # cant update to v2 on server
library(dplyr)
library(ggplot2)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
library(tictoc)
library(ncmeta)
library(parallel)
library(readr)
library(exactextractr)
library(tidyr)
library(data.table)
library(terra)
dir <- substr(getwd(), 1,2)

pData <- ifelse(dir == "C:", 
               "P:/41201785_okologisk_tilstand_2022_2023/data/",
               "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/")

```{r, echo=F}
Ecosystem <- "All" # e.g. "Skog og fjell"
Egenskap  <- "Abiotiske egenskaper" # e.g. "Primærproduksjon"
ECT       <- "Physical state characteristics" # e.g. "Structural state characteristic"
Contact   <- "Anders L. Kolstad" # e.g. "Anders Kolstad"
```

@@

```{r}
reg <- st_intersection(reg, nor)
```

```{r, fig.cap="Five accounting areas (regions) in Norway."}

tm_shape(reg) +
  tm_polygons(col="region")

@@

This takes about 20 sec per file/year, or 22 min on total. That is not too bad. About 6000 raster are read into memory. Here's a test for the effect of splitting over more cores.

```{r, eval=F}
write_stars(real_temp_summer, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp.tiff")
```

```{r, include=F}
summer_median_temp <- read_stars(paste0(pData, "climate_indicators/aggregated_climate_time_series/temp_median_summer.tiff"),
                                 proxy=F)
```

@@

```{r}
summer_median_temp <- summer_median_temp %>% 
  st_set_dimensions(names = c("x", "y", "v_YEAR")) %>%
  setNames("temperature")
```

```{r, fig.cap="Showing three random slices of the year dimension."}
ggplot()+
  geom_stars(data = summer_median_temp[,,,c(1,11,22)], downsample = c(10, 10, 0)) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))+
  facet_wrap(~v_YEAR)
```

@@

@@

```{r, eval=F}
saveRDS(summer_median_temp_ref, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp_ref.RData")
```

```{r, eval=T, include=F}
summer_median_temp_ref <- readRDS(paste0(pData, "climate_indicators/aggregated_climate_time_series/temp_median_summer_ref.RData"))
```

Pivot and turn dimension into attributes, and rename attribute:

```{r, eval=T}
summer_median_temp_ref_long <- summer_median_temp_ref %>% 
  split("median_sd") %>%
  setNames(c("reference_upper", "sd"))
```

@@

```{r}
y_var <- summer_median_temp %>%
  split(3) %>%
  c(summer_median_temp_ref_long)
```

@@

```{r, eval=F}
gc()
saveRDS(y_var_norm, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/summer_median_temp_normalised.RData")
```

```{r, include=FALSE}
y_var_norm <- readRDS("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/aggregated_climate_time_series/temp_median_summer_normalised.RData")
```
